---
title: "Simple plotting of NZ coastline"
author: "Adam Smith"
format: 
  html:
    toc: true
---

The purpose of this page is to show the easiest and best way I've found to access and plot New Zealand coastline. 

## Set up

```{r}
#| output: false
library(sf)
library(sfheaders)
library(southernMaps)
library(tidyverse)
theme_set(theme_bw())
```

## Load NZ coastline data

The coastline data come from the `southernMaps` package. Don't use `nzHires`; it's bad in many ways. 

### New Zealand Transverse Mercator

Here's how to access the coastline in NZTM (a nice coordinate system for NZ).

```{r}
nztm <- nzAndIslands |> st_as_sf()
```

### Make a version in WGS84

```{r}
nz84 <- nztm |> 
  st_transform(crs = 4326) # nztm is crs 2193
```

```{r}
#| eval: false
#| include: false
# Save
# save(nztm, file = 'data/nztm.Rdata')
# save(nz84, file = 'data/nz84.rda')

# Load previously saved version
# load(file = 'data/nztm.Rdata')
```

## Plot NZ Coastline

### Basic NZ plot

```{r}
nzp <- nztm |> 
  ggplot() +
  geom_sf(fill='grey', colour=0) 

nzp
```

### Hauraki Gulf

```{r}
nzhg <- nzp +
  xlim(1.75e6,1.82e6) + # longitude limits in NZTM
  ylim(5.91e6,6.0e6)    # latitude limits in NZTM

nzhg
```

### Fiordland

```{r}
nzfi <- nzp +
  xlim(1.09e6,  1.2e6) +
  ylim(4.87e6, 5.06e6)

nzfi
```
